--
title: "mIVD_mechanics"
author: "Jian"
date: '2024-12-15'
output: html_document
---

# Note: This script include the code used for analyzing mechanical test data of mouse FSU
# The input data are accessible at https://data.mendeley.com/datasets/9fxgx59jd7/1

## 1. Preparations
### 1.1 prepare the envs and path
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=T, warning=T, cache = FALSE, cache.lazy = FALSE, dev = "png",dpi = 300)
Sys.setenv(LANGUAGE = "en") 
options(stringsAsFactors = F, future.fork.enable = TRUE, future.globals.maxSize = 2000 * 1024^1024, 
        repos = c(CRAN="https://mirrors.tuna.tsinghua.edu.cn/CRAN/"),
        BioC_mirror="https://mirrors.tuna.tsinghua.edu.cn/bioconductor/")
```


### 1.2 clean the envs and set the data path
```{r}
rm(list = ls())
gc()
knitr::opts_knit$set(root.dir = "~/project/mouse_IVD/mechanics")
```



## 2. ultimate compression
### 2.1 prepare data
```{r}
data_fsu <- read.csv("./ultimate_compression_data.csv", row.names = 1)
```


### 2.2 all curve visulization
```{r, fig.width=8, fig.height=4}
library(ggplot2)
data_fsu <- na.omit(data_fsu)
p1 <- ggplot(data_fsu, aes(strain, stress, color = sample)) +
  geom_line(aes(linetype = group))+
  facet_wrap(~batch, scales = "free")+
  # scale_color_brewer(palette = "Set2")+
  theme_classic() + labs(title = "Stress-Strain curve")
p1
```


### 2.3 fit Force-strain curve
```{r, fig.width=12, fig.height=12}
# Note: The standarized curves for each sample are retained 
## The data after the ultimate point are removed
## To calculate the slopes and derivates, the curves should be fitted or smoothed firstly


# 1. smooth the curve using l-m model
data_fsu_lm <- data.frame()
for (s in unique(data_fsu$sample)) {
  fsu <- subset(data_fsu, sample == s)
  set.seed(123)
  fit <- loess(stress ~ strain, fsu, span=0.05, degree = 1)
  newx <- seq(0, 0.5, length.out = 1001)
  newy <- predict(fit, newdata = data.frame(strain = newx))
  fsu_new <- data.frame(strain = newx, stress = newy)
  fsu_new$sample <- s
  fsu_new$group <- unique(fsu$group)
  data_fsu_lm <- rbind(data_fsu_lm, fsu_new) 
}

data_fsu_lm <- na.omit(data_fsu_lm)



# 2. show the representative curves
# S3_0322 type A; S3_0418 type C
data_tmp <- subset(data_fsu_lm, sample %in% c("S3_0322","S3_0418"))

p2 <- ggplot(data_tmp, aes(strain, stress, color = group))+
  geom_line(linewidth =1)+
  scale_color_manual(values = c('#a5dff9','#FFBC42'))+ 
  theme_classic() + labs(title = "", y = "Stress (KPa)", x = "Strain (%)")
p2


#3. The data after the yield point are removed
data_fsu_lm1 <- data.frame()
for (s in unique(data_fsu_lm$sample)) {
  fsu <- subset(data_fsu_lm, sample == s)
  fsu <- fsu %>%
  mutate(
    slope = c(NA, diff(stress)/diff(strain)),  
    )
  fsu <- na.omit(fsu)
  first_negative_index <- which(fsu$slope < 0)
  if (isEmpty(first_negative_index)) {
    fsu <- fsu
  } else {
    first_negative_index <- first_negative_index[1] - 1
    fsu <- fsu[1:first_negative_index, ]
  }
  data_fsu_lm1 <- rbind(data_fsu_lm1, fsu)
}

p3 <- ggplot(data_fsu_lm1, aes(strain, stress, color = sample)) +
  geom_line(aes(linetype = group))+
  # facet_wrap(~sample, nrow = 3)+
  # scale_color_brewer(palette = "Set2")+
  theme_classic() + labs(title = "Stress-Strain curve")

sample_labels <- data_fsu_lm1 %>% group_by(sample) %>%
  filter(strain == max(strain))

p3 <- p3 + 
  geom_text_repel(data = sample_labels, 
            aes(label = sample, x = strain, y = stress),
            # position = position_stack(vjust = -0.5),
            fontface = "bold", size = 1)
p3
```


### 2.4 calculate yield, transition points and modules
```{r}
# After smoothing/fitting, the features for each sample curves can be calculated using segmented function 
library(segmented)

## 1. prepare each parameters
samples <- c()
yield_strain <- c()
yield_stress <- c()
trans_strain <- c()
Rsq_adj_val <- c()
module_toe <- c()
module_linear <- c()

for (s in unique(data_fsu_lm1$sample)) {
 fsu <- subset(data_fsu_lm1, sample == s) 
 yield_point <- fsu[which(fsu$stress == max(fsu$stress)),]$strain[1] # the yield point is the value of the maximum stress
 yield_force <- max(fsu$stress) # the maxium force
 fsu_linear <- fsu[fsu$strain < yield_point, ] # the endpoint of the toe and linear region
 
 ## 1. segment the curve into toe and linear region
 fsu_lm <- lm(stress~strain, data = fsu_linear)
 
 ## 2. find the transition point
 fsu_seg <- segmented(fsu_lm, seg.Z = ~strain, npsi = 1)
 trans_point <- fsu_seg$psi[2]
 Rsq_adj <- summary(fsu_seg)[["adj.r.squared"]] # test if it is significant
 
 ## 3. split the curve
 ### 3.1 toe region
 model_toe <- lm(stress~strain, data = fsu_linear[fsu_linear$strain < trans_point, ])
 slope_toe <- coef(model_toe)["strain"] ## module of toe region
 ### 3.2 linear region
 model_linear <- lm(stress~strain, data = fsu_linear[fsu_linear$strain > trans_point, ])
 slope_linear <- coef(model_linear)["strain"] ## module of linear region
 
 
 ## 4. combine the data
 samples <- c(samples, s)
 yield_strain <- c(yield_strain, yield_point)
 yield_stress <- c(yield_stress, yield_force)
 trans_strain <- c(trans_strain, trans_point)
 Rsq_adj_val <- c(Rsq_adj_val, Rsq_adj)
 module_toe <- c(module_toe, slope_toe)
 module_linear <- c(module_linear, slope_linear)
}


fsu_results <- data.frame(samples, yield_strain,yield_stress, trans_strain, Rsq_adj_val, module_toe, module_linear)
```



### 2.5 compare the differences
```{r, fig.height=3, fig.width=10}
library(ggpubr)

# 1. add the group information
fsu_results$group <-  plyr::mapvalues(fsu_results$sample, data_fsu_lm1$sample, data_fsu_lm1$group)

fsu_df <- reshape2::melt(fsu_results, id = c("samples", "group"))
fsu_df$group <- factor(fsu_df$group, levels = c("A", "C"))
table(fsu_results$group)

# 2. strain
p_fsu1 <- ggplot(subset(fsu_df, variable %in% c("trans_strain","yield_strain")), aes(group, value, fill=group))+
  geom_bar(stat="summary",fun=mean,position="dodge", width = 0.8)+ 
  geom_jitter(width = 0.3,size=5, pch=21, color="black", fill = "white")+
  stat_summary(geom = "errorbar", fun.data = 'mean_se', width = 0.3)+
  stat_compare_means(aes(x=group, y=value), method = "wilcox",
                     comparisons = list(c("A", "C")),
                     label = "p.format" )+
  # stat_compare_means(label.y = 0.9)+
  theme_classic()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())+ 
  scale_fill_manual(values = c('#a5dff9','#FFBC42')) +
  facet_wrap(~variable, nrow = 1)
p_fsu1

# 3. module 
p_fsu2 <- ggplot(subset(fsu_df, variable %in% c("module_toe","module_linear")), aes(group, value, fill=group))+
  geom_bar(stat="summary",fun=mean,position="dodge", width = 0.8)+ 
  geom_jitter(width = 0.3,size=5, pch=21, color="black", fill = "white")+
  stat_summary(geom = "errorbar", fun.data = 'mean_se', width = 0.3)+
  stat_compare_means(aes(x=group, y=value), method = "wilcox",
                     comparisons = list(c("A", "C")),
                     label = "p.format" )+
  # stat_compare_means(label.y = 0.9)+
  theme_classic()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())+ 
  scale_fill_manual(values = c('#a5dff9','#FFBC42')) +
  facet_wrap(~variable, nrow = 1)
p_fsu2

# 4. maximum stress
p_fsu3 <- ggplot(subset(fsu_df, variable %in% c("yield_stress")), aes(group, value, fill=group))+
  geom_bar(stat="summary",fun=mean,position="dodge", width = 0.8)+ 
  geom_jitter(width = 0.3,size=5, pch=21, color="black", fill = "white")+
  stat_summary(geom = "errorbar", fun.data = 'mean_se', width = 0.3)+
  stat_compare_means(aes(x=group, y=value), method = "wilcox",
                     comparisons = list(c("A", "C")),
                     label = "p.format" )+
  # stat_compare_means(label.y = 0.9)+
  theme_classic()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())+ 
  scale_fill_manual(values = c('#a5dff9','#FFBC42')) +
  facet_wrap(~variable, nrow = 1)
p_fsu3

  
(p_fsu1 | p_fsu2 | p_fsu3) + plot_layout(widths = c(1,1,0.5))
```



## 3. ultimate 3-point bending
### 3.1 prepare data
```{r}
data_bend <- read.csv("./ultimate_3point_bending_data.csv", row.names = 1)
```

### 3.2 all curve visualization
```{r, fig.width=10, fig.height=10}
library(ggplot2)

sample_labels <- data_bend %>% group_by(sample) %>%
  filter(Displacement == max(Displacement))

p1 <- ggplot(data_bend, aes(Displacement, Force, color = sample)) +
  geom_line(aes(linetype = group))+
  facet_wrap(~group)+
  # scale_color_brewer(palette = "Set2")+
  theme_classic() + labs(title = "Force-Displacement curve")
p1 <- p1 + 
  geom_text_repel(data = sample_labels, 
            aes(label = sample, x = Displacement, y = Force),
            # position = position_stack(vjust = -0.5),
            fontface = "bold", size = 3)
p1
```

### 3.3 fit curve
```{r}

## 1. smooth using l-m model
data_bend_lm <- data.frame()

for (s in unique(data_bend$sample)) {
  fsu <- subset(data_bend, sample == s)
  set.seed(123)
  fit <- loess(Force ~ Displacement, fsu, span=0.05, degree = 1)
  newx <- seq(0, 1.8, length.out = 1001)
  newy <- predict(fit, newdata = data.frame(Displacement = newx))
  fsu_new <- data.frame(Displacement = newx, Force = newy)
  fsu_new$sample <- s
  fsu_new$group <- unique(fsu$group)
  data_bend_lm <- rbind(data_bend_lm, fsu_new) 
}

data_bend_lm <- na.omit(data_bend_lm)

p3 <- ggplot(data_bend_lm, aes(Displacement, Force, color = sample)) +
  geom_line(aes(linetype = group))+
  # facet_wrap(~sample, nrow = 3)+
  # scale_color_brewer(palette = "Set2")+
  theme_classic() + labs(title = "Force-Displacement curve")

sample_labels <- data_bend_lm %>% group_by(sample) %>%
  filter(Displacement == max(Displacement))

p3 <- p3 + 
  geom_text_repel(data = sample_labels, 
            aes(label = sample, x = Displacement, y = Force),
            # position = position_stack(vjust = -0.5),
            fontface = "bold")
p3


## 2. show the representative curves
data_tmp <- subset(data_bend_lm, sample %in% c("M8_C7", "M21_C5","M18_C5"))

p2 <- ggplot(data_tmp, aes(Displacement, Force, color = group))+
  geom_line(linewidth =1)+
  scale_color_manual(values = c('#a5dff9',"#D2CE9E",'#FFBC42'))+ 
  theme_classic() + labs(title = "", y = "Force (N)", x = "Displacement (mm)")
p2

## 3. remove the data after the yield point
data_bend_lm1 <- data.frame()
for (s in unique(data_bend_lm$sample)) {
  fsu <- subset(data_bend_lm, sample == s)
  max_index <- which.max(fsu$Force)
  fsu <- fsu[1:max_index, ]
  data_bend_lm1 <- rbind(data_bend_lm1, fsu)
}
```

### 3.4 calculate yield, transition points and modules
```{r}
library(segmented)

samples <- c()
yield_strain <- c()
yield_stress <- c()
trans_strain <- c()
Rsq_adj_val <- c()
module_toe <- c()
module_linear <- c()

for (s in unique(data_bend_lm1$sample)) {
 fsu <- subset(data_bend_lm1, sample == s) 
 yield_point <- fsu[which(fsu$Force == max(fsu$Force)),]$Displacement[1]
 yield_force <- max(fsu$Force)
 fsu_linear <- fsu[fsu$Displacement < yield_point, ]
 
 fsu_lm <- lm(Force~Displacement, data = fsu_linear)
 fsu_seg <- segmented(fsu_lm, seg.Z = ~Displacement, npsi = 1)
 trans_point <- fsu_seg$psi[2]
 Rsq_adj <- summary(fsu_seg)[["adj.r.squared"]]
 
 model_toe <- lm(Force~Displacement, data = fsu_linear[fsu_linear$Displacement < trans_point, ])
 slope_toe <- coef(model_toe)["Displacement"]
 model_linear <- lm(Force~Displacement, data = fsu_linear[fsu_linear$Displacement > trans_point, ])
 slope_linear <- coef(model_linear)["Displacement"]
 
 samples <- c(samples, s)
 yield_strain <- c(yield_strain, yield_point)
 yield_stress <- c(yield_stress, yield_force)
 trans_strain <- c(trans_strain, trans_point)
 Rsq_adj_val <- c(Rsq_adj_val, Rsq_adj)
 module_toe <- c(module_toe, slope_toe)
 module_linear <- c(module_linear, slope_linear)
}

fsu_results_bending <- data.frame(samples, yield_strain,yield_stress, trans_strain, Rsq_adj_val, module_toe, module_linear)
```


### 3.5 compare the differences
```{r, fig.height=3, fig.width=10}
library(ggpubr)
# 1. Pmax: the property to anti-bending
data_Fmax <- data_bend %>% group_by(sample, group) %>% dplyr::summarise(max(Force))
colnames(data_Fmax) <- c("sample", "group", "Fmax")
data_Fmax$diameter <- plyr::mapvalues(data_Fmax$sample, data_bend$sample, data_bend$Diameter)
data_Fmax$diameter <- as.numeric(data_Fmax$diameter)
data_Fmax$Property <- data_Fmax$Fmax * 8 / (pi * (data_Fmax$diameter^3))

p_Pmax <- ggplot(data_Fmax, aes(group, Property, fill=group))+
  geom_bar(stat="summary",fun=mean,position="dodge", width = 0.8)+ 
  geom_jitter(width = 0.3,size=5, pch=21, color="black", fill = "white")+
  stat_summary(geom = "errorbar", fun.data = 'mean_se', width = 0.3)+
  stat_compare_means(aes(x=group, y=value), method = "t.test",
                     comparisons = list(c("AA", "AC"),c("AA", "CC"),c("AC", "CC")),
                     label = "p.format" )+
  # stat_compare_means(label.y = 0.9)+
  theme_classic()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())+ 
  labs(y = "Bending strength (N/mm2)")+
  scale_fill_manual(values = c('#a5dff9',"#D2CE9E",'#FFBC42'))

p_Pmax

# other properties
# 1. add the group information
fsu_results_bending$group <-  plyr::mapvalues(fsu_results_bending$sample, data_bend_lm1$sample, data_bend_lm1$group)
fsu_results_bending$diameter <-  data_Fmax$diameter

# 2. calculate the modules
## Note: The dFSU is approximated as cylinder
## 2.1 toe module
fsu_results_bending$module_toe_1 <- fsu_results_bending$module_toe*8^3/(48*(pi*(fsu_results_bending$diameter)^4/64)) 
## 2.2 linear module
fsu_results_bending$module_linear_1 <- fsu_results_bending$module_linear*8^3/(48*(pi*(fsu_results_bending$diameter)^4/64))

# 3. visualization
fsu_df <- reshape2::melt(fsu_results_bending, id = c("samples", "group"))
fsu_df$group <- factor(fsu_df$group, levels = c("AA","AC", "CC"))
table(fsu_results_bending$group)

# 4. strain
p_fsu1 <- ggplot(subset(fsu_df, variable %in% c("trans_strain","yield_strain")), aes(group, value, fill=group))+
  geom_bar(stat="summary",fun=mean,position="dodge", width = 0.8)+ 
  geom_jitter(width = 0.3,size=5, pch=21, color="black", fill = "white")+
  stat_summary(geom = "errorbar", fun.data = 'mean_se', width = 0.3)+
  stat_compare_means(aes(x=group, y=value), method = "t.test",
                     comparisons = list(c("AA", "AC"),c("AA", "CC"),c("AC", "CC")),
                     label = "p.format" )+
  # stat_compare_means(label.y = 0.9)+
  theme_classic()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())+ 
  scale_fill_manual(values = c('#a5dff9',"#D2CE9E",'#FFBC42'))+
  facet_wrap(~variable, nrow = 1)
p_fsu1

p_fsu2 <- ggplot(subset(fsu_df, variable %in% c("module_toe_1","module_linear_1")), aes(group, value, fill=group))+
  geom_bar(stat="summary",fun=mean,position="dodge", width = 0.8)+ 
  geom_jitter(width = 0.3,size=5, pch=21, color="black", fill = "white")+
  stat_summary(geom = "errorbar", fun.data = 'mean_se', width = 0.3)+
  stat_compare_means(aes(x=group, y=value), method = "t.test",
                     comparisons = list(c("AA", "AC"),c("AA", "CC"),c("AC", "CC")),
                     label = "p.format" )+
  # stat_compare_means(label.y = 0.9)+
  theme_classic()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())+ 
  scale_fill_manual(values = c('#a5dff9',"#D2CE9E",'#FFBC42'))+
  facet_wrap(~variable, nrow = 1)
p_fsu2


( p_Pmax| p_fsu2 | p_fsu1 ) + plot_layout(widths = c(0.5, 1,1))

```


## 4. Axial tension-compression testing
### 4.1 prepare data
```{r, fig.width=15, fig.height=15}
data_fsu <- read.csv("./cyclin_compression-tension_data.csv", row.names = 1)
```

### 4.2 split the waves
```{r, fig.width=15, fig.height=10}
# Note: Firstly, the cyclin curves are split into waves
# Then, each wave is split into compression and tension process
# The last wave is chose to compute properties for testing

data_fsu_final <- data.frame()
for (i in unique(data_fsu$sample)) {
  data_tmp <- subset(data_fsu, sample == i)
  data_tmp$force_ori <- data_tmp$force
  
  ## 1. using loess function to remove the curve noisy 
  data_tmp$force <- loess(data_tmp$force ~ seq_along(data_tmp$force), span=0.02, degree = 1)$fitted
  
  ## 2. calculating the derivatives
  wave_diff <- diff(sign(diff(data_tmp$force)))
  
  ## 3. determining the transition point for each wave
  peak_indices <- which(wave_diff < 0) + 1
  
  ## 4. splitting waves
  wave <- c(rep(paste0("wave_", 1), peak_indices[1]))
  for (k in 2:length(peak_indices)) {
    tmp <- rep(paste0("wave_", k), peak_indices[k] - peak_indices[k-1])
    wave <- c(wave, tmp)
    }
  wave <- c(wave, rep(paste0("wave_", length(peak_indices) + 1), length(data_tmp$force) - peak_indices[length(peak_indices)]))
  
  data_tmp$wave <- wave 
  
  ## 4. determine the process by derivatives
  data_tmp$wave_diff <- c(sign(diff(data_tmp$force))[1], sign(diff(data_tmp$force)))
  data_tmp$process <- "Compression"
  data_tmp[data_tmp$wave_diff > 0, ]$process <- "Tension"
  
  ## 5. filter waves with data points greater than 10
  library(dplyr)
  freq_table <- data_tmp %>% dplyr::count(wave)  
  valid_values <- freq_table %>% filter(n >= 10) %>% pull(wave)
  data_tmp <- data_tmp %>% filter(wave %in% valid_values) 
  
  ## 6. filter waves with incomplete compression process
  range_table <- subset(data_tmp, process == "Compression") %>% group_by(wave) %>% 
    dplyr::summarise(range = max(force) - min(force))  
  valid_values <- range_table %>% filter(range >= 0.8) %>% dplyr::pull(wave)
  data_tmp <- data_tmp %>% filter(wave %in% valid_values) 
  
  data_fsu_final <- rbind(data_fsu_final, data_tmp)
}


data_fsu_filter <- data_fsu_final

## 7. choose the last complete wave for each sample
data_fsu_filter$sample <- factor(data_fsu_filter$sample, levels = paste0("S", 1:20))
data_fsu_filter$wave <- gsub("wave_", "", data_fsu_filter$wave)
data_fsu_filter$wave <- as.numeric(data_fsu_filter$wave)
data_fsu_filter$sample_wave <- paste0(data_fsu_filter$sample, "_", data_fsu_filter$wave)

last_second_class_per_group <- data_fsu_filter %>%
  group_by(sample) %>%
  dplyr::arrange(wave) %>% 
  dplyr::mutate(class_rank = dense_rank(wave)) %>%  
  dplyr::filter(class_rank == length(unique(class_rank)) - 1) %>%  
  pull(sample_wave)  
data_fsu_filter <- subset(data_fsu_filter, sample_wave %in% as.character(last_second_class_per_group))

p1 <- ggplot(data_fsu_filter, aes(x = displacement, y = force_ori, col = process)) +
  geom_point(size = 0.5) +
  facet_wrap(~sample, scales = "free_x") +
  theme_bw() + labs(title = "force-displacement curve")
p1
```


### 4.3 define the NZ boundry, Stiffness, ROM
#### 4.3.1 compression
```{r}
NZ_length_compression <- c()
NZ_stiff_compression <- c()
Stiff_compression <- c()
Axis_ROM_compression <- c()
p_list <- list()

for (s in levels(data_fsu_filter$sample)) {
  ## 1. filter the process and sample for single curve
  data_nz <- subset(data_fsu_filter, process == "Compression")
  data_nz <- subset(data_nz, sample == s)
  
  ## 2. prepare the force-displacement data
  data_input <- data_nz[, c("displacement", "force_ori")]
  data_input$displacement <- data_input$displacement -  min(data_input$displacement)
  data_input$force_ori <- data_input$force_ori -  min(data_input$force_ori)
  rownames(data_input) <- 1:nrow(data_input)
  
  ## 3. double sigmoid fit
  ### 3.1 define the function
  double_sigmoid <- function(x, L1, k1, x01, L2, k2, x02) {
    L1 / (1 + exp(-k1 * (x - x01))) + L2 / (1 + exp(-k2 * (x - x02)))  
    }  

  ### 3.2 data for fit
  data_input <- data_nz[, c("displacement", "force_ori")]
  x <- data_input$force_ori - min(data_input$force_ori)
  y <- data_input$displacement - min(data_input$displacement)
  
  ## 3.3. prepare initial params  
  start_values <- list(L1 = max(y), k1 = 100, x01 = median(x), L2 = max(y), k2 = -0.1, x02 = median(x)*2)  
  
  ## 3.3 do nls fitting 
  relaxed_fit <- try(nls(y ~ double_sigmoid(x, L1, k1, x01, L2, k2, x02), 
                         start = start_values,  algorithm = "port",   
                         control = list(maxiter = 1000, warnOnly = TRUE),  
                         lower = c(0, -Inf, -Inf, 0, -Inf, -Inf),   
                         upper = c(Inf, Inf, Inf, Inf, Inf, Inf)))  
  
  ## 3.4. function and output
   if (!inherits(relaxed_fit, "try-error")) {  
     y_fit <- predict(relaxed_fit, newdata = list(x = x))
     # plot(x, y, main="Double Sigmoid Fit", xlab="Axis Load", ylab="Deflection", pch=19)  
     # lines(x, y_fit, col="red", lwd=2)  
     # legend("topleft", legend=c("Data", "Fit"),    
     #        col=c("black", "red"), lty=c(0,1), pch=c(19,NA))  
     } else {  
       cat("Fitting with relaxed constraints failed. Please check your data or initial guesses.\n")  
     }
  
  # 4. find the inflection_points
  library(pracma)
  x_fit <- seq(min(x), max(x), length.out = 1000)  
  y_fit <- predict(relaxed_fit, newdata = data.frame(x = x_fit))  
  df <- data.frame(x = x_fit, y = y_fit)  
  df$dy_dx <- gradient(y_fit, x_fit)  # 1st derivative
  df$d2y_dx2 <- gradient(df$dy_dx, x_fit)  # 2nd derivative 

  ## 4.1 set 2nd derivates = 0 as points  
  inflection_1st_y <- df$y[which.max(df$d2y_dx2)]  
  inflection_2nd_y <- df$y[which.min(df$d2y_dx2)]  

  inflection_1st_x <- df$x[which.max(df$d2y_dx2)]  
  inflection_2nd_x <- df$x[which.min(df$d2y_dx2)] 
  
  ## 4.2 plot
  p_list[[s]] <- ggplot(df) + 
    geom_point(aes(x = x, y = y), size = 0.1)+
    geom_line(aes(x = x, y = (dy_dx)/10), col = "red") +
    geom_line(aes(x = x, y = (d2y_dx2)/100), col = "blue") +
    geom_hline(yintercept = inflection_1st_y,lty="dashed") + 
    geom_hline(yintercept = inflection_2nd_y,lty="dashed") +
    geom_vline(xintercept = inflection_1st_x,lty="dashed") + 
    geom_vline(xintercept = inflection_2nd_x,lty="dashed") +
    theme_bw() + 
    labs(x = "load", y  = "displacement", title = paste0("Compression curve of ", s))
  
  ## 5. NC length
  NZ_length <- inflection_2nd_y - inflection_1st_y
  NZ_length_compression <- c(NZ_length_compression, NZ_length)
  
  Stiff_NZ <- (inflection_2nd_x - inflection_1st_x)/(inflection_2nd_y - inflection_1st_y)
  NZ_stiff_compression <- c(NZ_stiff_compression, Stiff_NZ)
  
  ## 6. Compressive stiffness:  bottom 25%-75% of compressive curve
  x_med <- median(x_fit)
  x_range <- x_med - min(x_fit)
  stiff <- 0.5*(x_med - min(x_fit)) /(predict(relaxed_fit, newdata = data.frame(x = min(x_fit) + 0.75*x_range)) - predict(relaxed_fit, newdata = data.frame(x = min(x_fit) + 0.25*x_range)))
  Stiff_compression <- c(Stiff_compression, stiff)
  
  ## 7. Axis range of motion
  y_max <- predict(relaxed_fit, newdata = data.frame(x = x_med))
  y_min <- predict(relaxed_fit, newdata = data.frame(x = min(x_fit)))
  Axis_ROM <- y_max - y_min
  Axis_ROM_compression <- c(Axis_ROM_compression, Axis_ROM)

}


## integration
data_compression <- data.frame(row.names = levels(data_fsu_filter$sample),
                               NZ_length_compression = NZ_length_compression,
                               Axis_ROM_compression = Axis_ROM_compression, 
                               Stiff_compression = Stiff_compression,
                               NZ_stiff_compression = NZ_stiff_compression)

```

#### 4.3.2 tension
```{r}
NZ_length_tensile <- c()
NZ_stiff_tensile <- c()
Stiff_tensile <- c()
Axis_ROM_tensile <- c()
p_list <- list()

for (s in levels(data_fsu_filter$sample)) {
  ## 1. filter the process and sample for single curve
  data_nz <- subset(data_fsu_filter, process == "Tension")
  data_nz <- subset(data_nz, sample == s)
  
  ## 2. prepare the force-displacement dataa 
  data_input <- data_nz[, c("displacement", "force_ori")]
  data_input$displacement <- data_input$displacement -  min(data_input$displacement)
  data_input$force_ori <- data_input$force_ori -  min(data_input$force_ori)
  rownames(data_input) <- 1:nrow(data_input)
  
  ## 3. double sigmoid fit
  ### 3.1 define the function
  double_sigmoid <- function(x, L1, k1, x01, L2, k2, x02) {
    L1 / (1 + exp(-k1 * (x - x01))) + L2 / (1 + exp(-k2 * (x - x02)))  
    }  

  ### 3.2 data for fit
  data_input <- data_nz[, c("displacement", "force_ori")]
  x <- data_input$force_ori - min(data_input$force_ori)
  y <- data_input$displacement - min(data_input$displacement)
  
  ## 3.3. prepare initial params  
  start_values <- list(L1 = max(y), k1 = 100, x01 = median(x), L2 = max(y), k2 = -0.1, x02 = median(x)*2)  
  
  ## 3.3 do nls fitting 
  relaxed_fit <- try(nls(y ~ double_sigmoid(x, L1, k1, x01, L2, k2, x02), 
                         start = start_values,  algorithm = "port",   
                         control = list(maxiter = 1000, warnOnly = TRUE),  
                         lower = c(0, -Inf, -Inf, 0, -Inf, -Inf),   
                         upper = c(Inf, Inf, Inf, Inf, Inf, Inf)))  
  
  ## 3.4. function and output
   if (!inherits(relaxed_fit, "try-error")) {  
     y_fit <- predict(relaxed_fit, newdata = list(x = x))
     # plot(x, y, main="Double Sigmoid Fit", xlab="Axis Load", ylab="Deflection", pch=19)  
     # lines(x, y_fit, col="red", lwd=2)  
     # legend("topleft", legend=c("Data", "Fit"),    
     #        col=c("black", "red"), lty=c(0,1), pch=c(19,NA))  
     } else {  
       cat("Fitting with relaxed constraints failed. Please check your data or initial guesses.\n")  
     }
  
  # 4. find the inflection_points
  library(pracma)
  x_fit <- seq(min(x), max(x), length.out = 1000)  
  y_fit <- predict(relaxed_fit, newdata = data.frame(x = x_fit))  
  df <- data.frame(x = x_fit, y = y_fit)  
  df$dy_dx <- gradient(y_fit, x_fit)  # 1st derivative
  df$d2y_dx2 <- gradient(df$dy_dx, x_fit)  # 2nd derivative 

  ## 4.1 set 2nd derivates = 0 as points  
  inflection_1st_y <- df$y[which.max(df$d2y_dx2)]  
  inflection_2nd_y <- df$y[which.min(df$d2y_dx2)]  

  inflection_1st_x <- df$x[which.max(df$d2y_dx2)]  
  inflection_2nd_x <- df$x[which.min(df$d2y_dx2)] 
  
  ## 4.2 plot
  p_list[[s]] <- ggplot(df) + 
    geom_point(aes(x = x, y = y), size = 0.1)+
    geom_line(aes(x = x, y = (dy_dx)/10), col = "red") +
    geom_line(aes(x = x, y = (d2y_dx2)/100), col = "blue") +
    geom_hline(yintercept = inflection_1st_y,lty="dashed") + 
    geom_hline(yintercept = inflection_2nd_y,lty="dashed") +
    geom_vline(xintercept = inflection_1st_x,lty="dashed") + 
    geom_vline(xintercept = inflection_2nd_x,lty="dashed") +
    theme_bw() + 
    labs(x = "load", y  = "displacement", title = paste0("Tensile curve of ", s))
  
  ## 5. NC length
  NZ_length <- inflection_2nd_y - inflection_1st_y
  NZ_length_tensile <- c(NZ_length_tensile, NZ_length)
  
  Stiff_NZ <- (inflection_2nd_x - inflection_1st_x)/(inflection_2nd_y - inflection_1st_y)
  NZ_stiff_tensile <- c(NZ_stiff_tensile, Stiff_NZ)
  
  ## 6. Tensile stiffness: top 25% - 75% of tensile curve
  x_med <- median(x_fit)
  x_range <- max(x_fit) - x_med 
  stiff <- 0.5*(max(x_fit)- x_med) /(predict(relaxed_fit, newdata = data.frame(x = x_med + 0.75*x_range)) - predict(relaxed_fit, newdata = data.frame(x = x_med + 0.25*x_range)))
  Stiff_tensile<- c(Stiff_tensile, stiff)
  
  ## 7. Axis range of motion
  y_max <- predict(relaxed_fit, newdata = data.frame(x = max(x_fit)))
  y_min <- predict(relaxed_fit, newdata = data.frame(x = x_med))
  Axis_ROM <- y_max - y_min
  Axis_ROM_tensile <- c(Axis_ROM_tensile, Axis_ROM)

}


## integration
data_tensile <- data.frame(row.names = levels(data_fsu_filter$sample),
                           NZ_length_tensile= NZ_length_tensile,
                           Axis_ROM_tensile = Axis_ROM_tensile, 
                           Stiff_tensile = Stiff_tensile,
                           NZ_stiff_tensile = NZ_stiff_tensile)

```

### 4.4 compare the differences
```{r, fig.width=16, fig.height=3}
data_cyclin <- cbind(data_compression, data_tensile)
data_cyclin$disc_type = mapvalues(levels(data_fsu_filter$sample), data_fsu_filter$sample, data_fsu_filter$type)
data_cyclin$disc_height = mapvalues(levels(data_fsu_filter$sample), data_fsu_filter$sample, data_fsu_filter$height)
data_cyclin$disc_area = mapvalues(levels(data_fsu_filter$sample), data_fsu_filter$sample, data_fsu_filter$area)

data_cyclin$disc_height <- as.numeric(data_cyclin$disc_height)
data_cyclin$disc_area <- as.numeric(data_cyclin$disc_area)

## normalization
data_cyclin$NZ_length_compression <- data_cyclin$NZ_length_compression/data_cyclin$disc_height
data_cyclin$Axis_ROM_compression <- data_cyclin$Axis_ROM_compression/data_cyclin$disc_height
data_cyclin$Stiff_compression <- data_cyclin$Stiff_compression/data_cyclin$disc_area*data_cyclin$disc_height
data_cyclin$NZ_stiff_compression <- data_cyclin$NZ_stiff_compression/data_cyclin$disc_area*data_cyclin$disc_height

data_cyclin$NZ_length_tensile <- data_cyclin$NZ_length_tensile/data_cyclin$disc_height
data_cyclin$Axis_ROM_tensile <- data_cyclin$Axis_ROM_tensile/data_cyclin$disc_height
data_cyclin$Stiff_tensile <- data_cyclin$Stiff_tensile/data_cyclin$disc_area*data_cyclin$disc_height
data_cyclin$NZ_stiff_tensile <- data_cyclin$NZ_stiff_tensile/data_cyclin$disc_area*data_cyclin$disc_height

## visualization
fsu_df <- reshape2::melt(data_cyclin[, 1:9],id = "disc_type")
colnames(fsu_df) <- c("group", "variable", "value")
fsu_df$value <- as.numeric(fsu_df$value)

p_fsu1 <- ggplot(subset(fsu_df, variable %in% c("NZ_length_compression","NZ_length_tensile")), aes(group, value, fill=group))+
  geom_bar(stat="summary",fun=mean,position="dodge", width = 0.8)+ 
  geom_jitter(width = 0.3,size=5, pch=21, color="black", fill = "white")+
  stat_summary(geom = "errorbar", fun.data = 'mean_se', width = 0.3)+
  stat_compare_means(aes(x=group, y=value), method = "wilcox",
                     comparisons = list(c("A", "C")),
                     label = "p.format" )+
  # stat_compare_means(label.y = 0.9)+
  theme_classic()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())+ 
  scale_fill_manual(values = c('#a5dff9','#FFBC42')) +
  facet_wrap(~variable, nrow = 1)
p_fsu1

p_fsu2 <- ggplot(subset(fsu_df, variable %in% c("Axis_ROM_compression","Axis_ROM_tensile")), aes(group, value, fill=group))+
  geom_bar(stat="summary",fun=mean,position="dodge", width = 0.8)+ 
  geom_jitter(width = 0.3,size=5, pch=21, color="black", fill = "white")+
  stat_summary(geom = "errorbar", fun.data = 'mean_se', width = 0.3)+
  stat_compare_means(aes(x=group, y=value), method = "wilcox",
                     comparisons = list(c("A", "C")),
                     label = "p.format" )+
  # stat_compare_means(label.y = 0.9)+
  theme_classic()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())+ 
  scale_fill_manual(values = c('#a5dff9','#FFBC42')) +
  facet_wrap(~variable, nrow = 1)
p_fsu2

p_fsu3 <- ggplot(subset(fsu_df, variable %in% c("Stiff_compression", "Stiff_tensile")), aes(group, log2(value+1), fill=group))+
  geom_bar(stat="summary",fun=mean,position="dodge", width = 0.8)+ 
  geom_jitter(width = 0.3,size=5, pch=21, color="black", fill = "white")+
  stat_summary(geom = "errorbar", fun.data = 'mean_se', width = 0.3)+
  stat_compare_means(aes(x=group, y=value), method = "wilcox",
                     comparisons = list(c("A", "C")),
                     label = "p.format" )+
  # stat_compare_means(label.y = 0.9)+
  theme_classic()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())+ 
  scale_fill_manual(values = c('#a5dff9','#FFBC42')) +
  facet_wrap(~variable, nrow = 1)
p_fsu3


p_fsu4 <- ggplot(subset(fsu_df, variable %in% c("NZ_stiff_compression", "NZ_stiff_tensile")), aes(group, value, fill=group))+
  geom_bar(stat="summary",fun=mean,position="dodge", width = 0.8)+ 
  geom_jitter(width = 0.3,size=5, pch=21, color="black", fill = "white")+
  stat_summary(geom = "errorbar", fun.data = 'mean_se', width = 0.3)+
  stat_compare_means(aes(x=group, y=value), method = "wilcox",
                     comparisons = list(c("A", "C")),
                     label = "p.format" )+
  # stat_compare_means(label.y = 0.9)+
  theme_classic()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())+ 
  scale_fill_manual(values = c('#a5dff9','#FFBC42')) +
  facet_wrap(~variable, nrow = 1)
p_fsu4

  
p_fsu1 | p_fsu2 | p_fsu3 | p_fsu4
```




